---
title: "Hw 3"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(lubridate)
library(ggthemes)
```


```{r}
data = read_csv("county_data.csv")
wake = data |>
  filter(COUNTYFP=="183")
wake 

#Confirms the dataset has only Wake County residents
sum(wake$triangle_census_total_population)
```


What percentage of the Wake County population resides in food deserts?
```{r}
#Find the total number of people in a food desert and not in a food desert
wake |>
  group_by(food_des) |>
  summarize(triangle_census_total_population=sum(triangle_census_total_population)) |>
  ungroup()
```
`
```{r}
#Find the percentage using the number of people in a food desert and the total number of people in the county
#1 is not in a food desert, 0 is in a food desert
(906882 / 1069079) * 100
```


Evaluate the demographic makeup of the food deserts—are residents of food deserts more likely to be poor or not own cars than the residents of Wake County as a whole (again, make sure you only include Wake County residents)? Calculate the percentage of households that do not own a vehicle, and the percentage of households who are low income, in both the food desert and Wake County overall.

```{r}
#Total number of households that don't own a vehicle
wake |>
  group_by(food_des) |>
  summarize(triangle_census_zero_vehicle_households=sum(triangle_census_zero_vehicle_households)) |>
  ungroup()

#Total number of households in Wake County 
sum(wake$triangle_census_total_households) 
# value = 400172

#1 is not in a food desert, 0 is in a food desert
#Food desert households without a vehicle 
(11267 / 400172) * 100
# Value = 2.815539

#Not food desert households without a vehicle 
(4611 / 400172) * 100
# Value = 1.152255

```

```{r}
#Total number of households that are low income
wake |>
  group_by(food_des) |>
  summarize(triangle_census_households_income_less_than_35k = sum(triangle_census_households_income_less_than_35k)) |>
  ungroup()

#Total number of households in Wake County 
sum(wake$triangle_census_total_households) 
# value = 400172

#1 is not in a food desert, 0 is in a food desert
#Food desert households that are low income
(58519 / 400172) * 100
# Value = 14.62346

#Not food desert households that are low income
(16714	 / 400172) * 100
# Value = 4.176704

```

2 Mile Criterion 
```{r}
#Add new data with classified food desert block group utilizing two mile criterion
wake2 = read_csv("county_data2.csv") |>
  filter(COUNTYFP=="183")
wake2
```

```{r}
#Find the total number of people in a food desert and not in a food desert
wake2 |>
  group_by(twom_fd) |>
  summarize(triangle_census_total_population=sum(triangle_census_total_population)) |>
  ungroup()

#Find the total number of people in the county
sum(wake2$triangle_census_total_population)

#Find the percentage using the number of people in a food desert and the total number of people in the county
#1 is not in a food desert, 0 is in a food desert
(400285 / 1069079) * 100
```

```{r}
#Total number of households that don't own a vehicle
wake2 |>
  group_by(twom_fd) |>
  summarize(triangle_census_zero_vehicle_households=sum(triangle_census_zero_vehicle_households)) |>
  ungroup()

#Total number of households in Wake County 
sum(wake2$triangle_census_total_households) 
# value = 400172

#1 is not in a food desert, 0 is in a food desert
#Food desert households without a vehicle 
(3186 / 400172) * 100
# Value = 0.7961577

#Not food desert households without a vehicle 
(12692 / 400172) * 100
# Value = 3.171636

```

```{r}
#Total number of households that are low income
wake2 |>
  group_by(twom_fd) |>
  summarize(triangle_census_households_income_less_than_35k = sum(triangle_census_households_income_less_than_35k)) |>
  ungroup()

#Total number of households in Wake County 
sum(wake2$triangle_census_total_households) 
# value = 400172

#1 is not in a food desert, 0 is in a food desert
#Food desert households that are low income
(22228 / 400172) * 100
# Value = 5.554612

#Not food desert households that are low income
(53005	/ 400172) * 100
# Value = 13.24555

```




